\documentclass[a4paper]{article}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{mathtools}
\usepackage{soul}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage[T1]{fontenc}
\usepackage{etoolbox}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{array}
\usepackage{amsthm}
\usepackage{titlesec}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,calc}
\renewcommand{\baselinestretch}{1.2}

\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\graphicspath{{C:/Users/teoso/OneDrive/Documents/Tugas Kuliah/Template Math Depart/}}

\newtheorem{theorem}{Theorem}
\newtheorem*{teorema}{Teorema}
\newtheorem*{definisi}{Definisi}
\theoremstyle{definition}
\newtheorem*{bukti}{Bukti}

\newcommand{\Arg}{\text{Arg}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}

\begin{document}
\fancyhead[L]{\textit{Teosofi Hidayah Agung}}
\fancyhead[R]{\textit{5002221132}}
\pagestyle{fancy}
\section*{ETS Formal Logic}

\subsection*{Questions}
\begin{enumerate}
  \item Given the Boolean formula in CNF:
        \[ (\neg a + b + \neg c)(a + \neg c + d)(a + \neg c + \neg d)(a + c + d)(a + c + \neg d)(\neg b + c + d)(\neg a + b + c)(\neg a + \neg b + \neg c) \]
        Check the satisfiability of the above formula using any algorithm.

  \item Agent Alina is preparing a stealth operation. To succeed, she must choose exactly one disguise, one route, and one gadget from the following options:
        \begin{itemize}
          \item[(i)] Disguise: Doctor ($d_1$), Tourist ($d_2$), and Delivery Person ($d_3$)
          \item[(ii)] Route: Sewer Tunnel ($r_1$), Rooftops ($r_2$), and Market Streets ($r_3$)
          \item[(iii)] Gadget: Invisibility Cloak ($g_1$), Grappling Hook ($g_2$), and Translator Earpiece ($g_3$)
        \end{itemize}
        However, she has certain constraints based on intel from the headquarter.
        \begin{itemize}
          \item[(i)] If Alina picks the Doctor disguise, she cannot go through the Market Street.
          \item[(ii)] She must use either the Rooftops or the Invisibility Cloak.
          \item[(iii)] The Translator Earpiece is useless in the Sewer Tunnel.
          \item[(iv)] She can't wear a Tourist disguise if she uses the Grappling Hook.
        \end{itemize}

        \begin{enumerate}
          \item[a.] Express each constraint into a clause (using the introduced variables)
          \item[b.] Can Agent Alina complete the mission? Explain your answer and show the satisfiability assignment for each variable.
        \end{enumerate}
\end{enumerate}

\subsection*{Solutions}

\begin{enumerate}
  \item Let's use the resolution method to check the satisfiability. Let $+$ denote disjunction ($\lor$) and the multiplication between parentheses denote conjunction ($\land$).

        First, check the clauses $(a \lor \neg c \lor d)$ and $(a \lor \neg c \lor \neg d)$. By resolving with respect to the variable $d$, we obtain:
        \[ (a \lor \neg c) \]

        Second, check the clauses $(a \lor c \lor d)$ and $(a \lor c \lor \neg d)$. By resolving with respect to the variable $d$, we obtain:
        \[ (a \lor c) \]

        From the two resolution results above, namely $(a \lor \neg c)$ and $(a \lor c)$, resolving with respect to $c$ yields:
        \[ a \]
        This means the variable $a$ must evaluate to \textbf{True (1)}.

        Next, substitute $a = 1$ (so $\neg a = 0$) into the clauses containing $\neg a$:
        \begin{itemize}
          \item $(\neg a \lor b \lor \neg c) \implies (0 \lor b \lor \neg c) \implies (b \lor \neg c)$
          \item $(\neg a \lor b \lor c) \implies (0 \lor b \lor c) \implies (b \lor c)$
          \item $(\neg a \lor \neg b \lor \neg c) \implies (0 \lor \neg b \lor \neg c) \implies (\neg b \lor \neg c)$
        \end{itemize}

        Resolve the clauses $(b \lor \neg c)$ and $(b \lor c)$ to obtain:
        \[ b \]
        So, the variable $b$ must evaluate to \textbf{True (1)}.

        Substitute $b = 1$ (so $\neg b = 0$) into the clause $(\neg b \lor \neg c)$ above:
        \[ 0 \lor \neg c \implies \neg c \]
        Thus, the variable $c$ must evaluate to \textbf{False (0)}.

        Finally, consider the remaining clause from the original formula that has not been used, which is $(\neg b \lor c \lor d)$. Substitute $b = 1$ and $c = 0$:
        \[ 0 \lor 0 \lor d \implies d \]
        This means the variable $d$ must evaluate to \textbf{True (1)}.

        Since we can find a truth assignment that satisfies the formula ($a = 1, b = 1, c = 0, d = 1$) without leading to an empty clause (contradiction), we conclude that the formula is \textbf{Satisfiable}.

  \item \begin{enumerate}
          \item[a.] Expressing each constraint as a propositional clause:
                \begin{itemize}
                  \item[(i)] If Alina picks Doctor ($d_1$), she cannot go through Market Street ($r_3$): \\ $d_1 \to \neg r_3 \equiv \neg d_1 \lor \neg r_3$
                  \item[(ii)] She must use either Rooftops ($r_2$) or Invisibility Cloak ($g_1$): \\ $r_2 \lor g_1$
                  \item[(iii)] Translator Earpiece ($g_3$) is useless in Sewer Tunnel ($r_1$) -- meaning she cannot use both: \\ $\neg (g_3 \land r_1) \equiv \neg g_3 \lor \neg r_1$
                  \item[(iv)] If she uses Grappling Hook ($g_2$), she cannot wear Tourist disguise ($d_2$): \\ $g_2 \to \neg d_2 \equiv \neg g_2 \lor \neg d_2$
                \end{itemize}

          \item[b.] \textbf{Yes}, Agent Alina can complete the mission. Here is the proof by finding at least one valid truth assignment.

                From the main premise, Alina must choose \textbf{exactly one} Disguise, Route, and Gadget.
                Based on constraint (ii), she must have at least one of $r_2$ or $g_1$ set to True.

                Let's pick a scenario where $r_2 = 1$ (Rooftops) and $g_1 = 1$ (Invisibility Cloak). Consequently, the other options in their respective categories are eliminated:
                \begin{itemize}
                  \item Route: Because $r_2 = 1 \implies r_1 = 0 \text{ and } r_3 = 0$.
                  \item Gadget: Because $g_1 = 1 \implies g_2 = 0 \text{ and } g_3 = 0$.
                \end{itemize}

                Let's check if this assignment violates any other constraints:
                \begin{itemize}
                  \item Constraint (i): $\neg d_1 \lor \neg r_3 \implies \neg d_1 \lor 1 \equiv \mathbf{1}$ (\textit{Satisfied}, regardless of $d_1$).
                  \item Constraint (iii): $\neg g_3 \lor \neg r_1 \implies 1 \lor 1 \equiv \mathbf{1}$ (\textit{Satisfied}).
                  \item Constraint (iv): $\neg g_2 \lor \neg d_2 \implies 1 \lor \neg d_2 \equiv \mathbf{1}$ (\textit{Satisfied}, regardless of $d_2$).
                \end{itemize}

                Since all constraints are satisfied, Alina just needs to pick one disguise. Suppose she picks Doctor ($d_1 = 1$), which means $d_2 = 0$ and $d_3 = 0$.

                The satisfiability assignment model for the variables is:
                \begin{itemize}
                  \item \textbf{Disguise:} $d_1 = 1,\quad d_2 = 0,\quad d_3 = 0$
                  \item \textbf{Route:} $r_1 = 0,\quad r_2 = 1,\quad r_3 = 0$
                  \item \textbf{Gadget:} $g_1 = 1,\quad g_2 = 0,\quad g_3 = 0$
                \end{itemize}
                This combination is valid and satisfies all rules.
        \end{enumerate}
\end{enumerate}
\end{document}